Capstone Movielens Report

### 2025-12-16 reports/capstone-movielens.site/02.5.UME-model.Rmd
Report section documented: UMGYDE Model Tuning: Step 1 (`degree = 0`) / Pre-Configuration.

::: {.noteblock data-latex=""}
The functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UM_effect.cv](#func.regularize.test_lambda.UM_effect.cv) are described in the sections *Common Helper Functions /* [Regularization Functions](#appndx_a.CHF.regularization) and  *Models Training: Support Functions /* [UME Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The helper function [get_fine_tune.param.endpoints](#func.get_fine_tune.param.endpoints) used in the code snippet above is described in the sections *Common Helper Functions /* [ Regularization Functions](#appndx_a.CHF.regularization) of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The custom functions [model.tune.param_range](#func.model.tune.param_range) and [regularize.test_lambda.UM_effect.cv](#func.regularize.test_lambda.UM_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Regularization Functions](#appndx_a.CHF.regularization) and  *Models Training: Support Functions /* [UME Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

>>>>

::: {.noteblock data-latex=""}
The functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UM_effect.cv](#func.regularize.test_lambda.UM_effect.cv) are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UME Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The helper function [get_fine_tune.param.endpoints](#func.get_fine_tune.param.endpoints) used in the code snippet above is described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The custom functions [model.tune.param_range](#func.model.tune.param_range) and [regularize.test_lambda.UM_effect.cv](#func.regularize.test_lambda.UM_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UME Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

### 2025-12-16 reports/capstone-movielens.site/02.6.UMGE-model.Rmd
Report section documented: UMGYDE Model Tuning: Step 1 (`degree = 0`) / Pre-Configuration.

::: {.noteblock data-latex=""}
The functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UMG_effect.cv](#func.regularize.test_lambda.UMG_effect.cv) are described in the sections *Common Helper Functions /* [ Regularization Functions](#appndx_a.CHF.regularization) and  *Models Training: Support Functions /* [UMGE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The custom functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UMG_effect.cv](#func.regularize.test_lambda.UMG_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Regularization Functions](#appndx_a.CHF.regularization) and  *Models Training: Support Functions /* [UMGE Model: Regularization] of [Appendix A: Support Functions], respectively.
:::

@@

::: {.noteblock data-latex=""}
The helper function [get_fine_tune.param.endpoints](#func.get_fine_tune.param.endpoints) used in code snippet above is described in the sections *Common Helper Functions /* [ Regularization Functions](#appndx_a.CHF.regularization) of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The custom functions [model.tune.param_range](#func.model.tune.param_range) and [regularize.test_lambda.UMG_effect.cv](#func.regularize.test_lambda.UMG_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Regularization Functions](#appndx_a.CHF.regularization) and  *Models Training: Support Functions /* [UMGE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

>>>>

::: {.noteblock data-latex=""}
The functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UMG_effect.cv](#func.regularize.test_lambda.UMG_effect.cv) are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UMGE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The custom functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UMG_effect.cv](#func.regularize.test_lambda.UMG_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UMGE Model: Regularization] of [Appendix A: Support Functions], respectively.
:::

@@

::: {.noteblock data-latex=""}
The helper function [get_fine_tune.param.endpoints](#func.get_fine_tune.param.endpoints) used in code snippet above is described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The custom functions [model.tune.param_range](#func.model.tune.param_range) and [regularize.test_lambda.UMG_effect.cv](#func.regularize.test_lambda.UMG_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UMGE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

### 2025-12-16 reports/capstone-movielens.site/02.7.UMGYE-model.Rmd
Report section documented: UMGYDE Model Tuning: Step 1 (`degree = 0`) / Pre-Configuration.

::: {.noteblock data-latex=""}
The functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UMGY_effect.cv](#func.regularize.test_lambda.UMGY_effect.cv) are described in the sections *Common Helper Functions /* [ Regularization Functions](#appndx_a.CHF.regularization) and  *Models Training: Support Functions /* [UMGYE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The custom functions [model.tune.param_range](#func.model.tune.param_range) and [regularize.test_lambda.UMGY_effect.cv](#func.regularize.test_lambda.UMGY_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Regularization Functions](#appndx_a.CHF.regularization) and  *Models Training: Support Functions /* [UMGYE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

>>>>

::: {.noteblock data-latex=""}
The functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UMGY_effect.cv](#func.regularize.test_lambda.UMGY_effect.cv) are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UMGYE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

@@

::: {.noteblock data-latex=""}
The custom functions [model.tune.param_range](#func.model.tune.param_range) and [regularize.test_lambda.UMGY_effect.cv](#func.regularize.test_lambda.UMGY_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UMGYE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

### 2025-12-16 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd
Report section documented: UMGYDE Model Tuning: Step 1 (`degree = 0`) / Pre-Configuration.

  - [tune.model_param](#func.tune.model_param) and [RMSE_kable](#func.RMSE_kable) (described in the *Common Helper Functions /* [Regularization Functions](#appndx_a.CHF.regularization) section) to update and print the table above.

  - [train_UMGY_SmoothedDay_effect.RMSE.cv.degree0](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0) (described in the *Models Training: Support Functions /* [UMGYDE Model: Tuning `loess` Params] section) to calculate the *RMSE* for the *UMGE Model*;
:::

Below is the visual representation of the `span` parameter values range we have figured out in this step of tuning:
```{r}
plt.title = "UMGYDE Model, `loess` function args: range of `span` values for `degree = 0`"
lss.UMGYDE.preset.degree0.result$tuned.result |>
  data.plot(plt.title,
              xname = "parameter.value", 
              yname = "RMSE", 
              xlabel = "spans", 
              ylabel = "RMSE")
rm(plt.title)
```

>>>>

  - [tune.model_param](#func.tune.model_param) (described in the *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) section) to to select the best value of the *tuning parameter* from the given sequence of the *parameter values* (passed in the [param_values](#func.tune.model_param.args.param_values) argument).

  - [train_UMGY_SmoothedDay_effect.RMSE.cv.degree0](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0) (described in the *Models Training: Support Functions /* [UMGYDE Model: Tuning `loess` Params] section) used as an auxiliary function passed in the [fn_tune.test.param_value](#func.tune.model_param.args.fn_tune.test.param_value) argument of the [tune.model_param](#func.tune.model_param) to test the *UMGYDE Model* with the [loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) `degree` parameter set to $0$.

:::

Below is a visual representation of the dependence of the *RMSE* value on the [loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) `span` parameter value over the selected interval, with the parameter `degree = 0`, as we have figured out in this step of tuning:
```{r}
plt.title = "UMGYDE Model, `loess` function args: range of `span` values for `degree = 0`"
lss.UMGYDE.preset.degree0.result$tuned.result |>
  data.plot(plt.title,
              xname = "parameter.value", 
              yname = "RMSE", 
              xlabel = "spans", 
              ylabel = "RMSE")
rm(plt.title)
```

::: {.noteblock data-latex=""}
In the code snippet above, we use the custom data visualization function [data.plot](#func.data.plot) described in the *Data Helper Functions /* [Data Visualization Functions] section of [Appendix A](#appndx_a).
:::

### 2025-12-16 reports/capstone-movielens.site/05.appendix.Rmd
Report section documented: UMGYDE Model Tuning: Step 1 (`degree = 0`) / Pre-Configuration.

### Regularization Functions {#appndx_a.CHF.regularization}
\

@@

Selects the best values from a pair of matched vectors: a vector of sorted *regularization parameter* $\lambda$ values and a vector of corresponding *Root Mean Squared Errors (RMSE)* values passed in the [param_values](#func.tune.model_param.args.param_values) and [RMSEs](#func.get_best_param.result.args.RMSEs) arguments, respectively.

@@

::: { #func.get_best_param.result.args.param_values .sidebar }
  - **param_values:** Numeric vector containing a sorted list of the *regularization parameter* $\lambda$ values to search for the best one during the *Model Regularization* process;

@@

This is a *utility function* used to select the *best values* during the *Model Regularization* process, utilizing the *penalized approach* to regularize predictions as explained in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-2].

With this approach, the best value of the *regularization parameter* $\lambda$ (we call this the *best value of the regularization parameter* or the *best value of $\lambda$*) is considered to be the value corresponding to the minimum value of the *RMSE* for the model being tested.

@@

Numeric vector containing a pair of values: the best *regularization parameter* $\lambda$ value and corresponding *minimal RMSE*. 
:::

@@

@@

::: {.noteblock data-latex=""}
*Regularization* is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above this function is passed in the [fn_tune.test.param_value](#func.tune.model_param.args.fn_tune.test.param_value) argument to the [tune.model_param](#func.tune.model_param) function described above in the section *Common Helper Functions /* [Regularization Functions] of *this Appendix*.

@@

::: {.noteblock data-latex=""}
This function is an auxiliary function specifically designed to be passed in the `fn_tune.test.param_value` argument to the [tune.model_param](#func.tune.model_param) and [model.tune.param_range](#func.model.tune.param_range) functions described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
*Regularization* is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
::: {.noteblock data-latex=""}
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above the `train_UMGY_SmoothedDay_effect.RMSE.cv.degree0` function is passed in the [fn_tune.test.param_value](#func.tune.model_param.args.fn_tune.test.param_value) argument to the [tune.model_param](#func.tune.model_param) function described above in the section *Common Helper Functions /* [Regularization Functions] of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
This is an auxiliary function intended to be passed in the `fn_tune.test.param_value` argument to the [tune.model_param](#func.tune.model_param) and [model.tune.param_range](#func.model.tune.param_range) functions described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above the `train_UMGY_SmoothedDay_effect.RMSE.cv.degree1` function is passed in the [fn_tune.test.param_value](#func.model.tune.param_range.args.fn_tune.test.param_value) argument to the [model.tune.param_range](#func.model.tune.param_range) function described above in the section *Common Helper Functions /* [Regularization Functions] of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above the `train_UMGY_SmoothedDay_effect.RMSE.cv.degree2` function is passed in the [fn_tune.test.param_value](#func.model.tune.param_range.args.fn_tune.test.param_value) argument to the [model.tune.param_range](#func.model.tune.param_range) function described above in the section *Common Helper Functions /* [Regularization Functions] of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above this function is passed in the [fn_tune.test.param_value](#func.model.tune.param_range.args.fn_tune.test.param_value) argument to the [model.tune.param_range](#func.model.tune.param_range) function described above in the section *Common Helper Functions /* [Regularization Functions] of *this Appendix*.

>>>>

### Model Tuning Utils {#appndx_a.CHF.model_tuning_utils}
\

@@

Selects the best values from a pair of matched vectors: a vector of sorted *tuning parameter* values and a vector of corresponding *Root Mean Squared Errors (RMSE)* values passed in the [param_values](#func.tune.model_param.args.param_values) and [RMSEs](#func.get_best_param.result.args.RMSEs) arguments, respectively.

@@

::: { #func.get_best_param.result.args.param_values .sidebar }
  - **param_values:** Numeric vector containing a sorted list of the *tuning parameter* values to search for the best one during the *Model Regularization* process;

@@

This is a *utility function* used to select the *best values* during the *Model Parameter Tuning* process (including the *Model Regularization* procedure that utilizes the *penalized approach* to regularize predictions as explained in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-2]).

With this approach, the best value of the *tuning parameter* is considered to be the value corresponding to the minimum value of the *RMSE* for the model being tested.

@@

Numeric vector containing a pair of values: the best *tuning parameter* value and corresponding *minimal RMSE*. 
:::

@@

With the approach used in this algorithm, the best value of the *tuning parameter* is considered to be the value corresponding to the minimum value of the *Root Mean Square Error (RMSE)* for the model being tested.

@@

::: {.noteblock data-latex=""}
*Regularization* is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Model Tuning Utils] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above this function is passed in the [fn_tune.test.param_value](#func.tune.model_param.args.fn_tune.test.param_value) argument to the [tune.model_param](#func.tune.model_param) function described above in the section *Common Helper Functions /* [Model Tuning Utils] of *this Appendix*.

@@

::: {.noteblock data-latex=""}
This function is an auxiliary function specifically designed to be passed in the `fn_tune.test.param_value` argument to the [tune.model_param](#func.tune.model_param) and [model.tune.param_range](#func.model.tune.param_range) functions described above in the *Common Helper Functions /* [Model Tuning Utils] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
*Regularization* is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Model Tuning Utils] section of *this Appendix*.
::: {.noteblock data-latex=""}
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above the `train_UMGY_SmoothedDay_effect.RMSE.cv.degree0` function is passed in the [fn_tune.test.param_value](#func.tune.model_param.args.fn_tune.test.param_value) argument to the [tune.model_param](#func.tune.model_param) function described above in the section *Common Helper Functions /* [Model Tuning Utils] of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
This is an auxiliary function intended to be passed in the `fn_tune.test.param_value` argument to the [tune.model_param](#func.tune.model_param) and [model.tune.param_range](#func.model.tune.param_range) functions described above in the *Common Helper Functions /* [Model Tuning Utils] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above the `train_UMGY_SmoothedDay_effect.RMSE.cv.degree1` function is passed in the [fn_tune.test.param_value](#func.model.tune.param_range.args.fn_tune.test.param_value) argument to the [model.tune.param_range](#func.model.tune.param_range) function described above in the section *Common Helper Functions /* [Model Tuning Utils] of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above the `train_UMGY_SmoothedDay_effect.RMSE.cv.degree2` function is passed in the [fn_tune.test.param_value](#func.model.tune.param_range.args.fn_tune.test.param_value) argument to the [model.tune.param_range](#func.model.tune.param_range) function described above in the section *Common Helper Functions /* [Model Tuning Utils] of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above this function is passed in the [fn_tune.test.param_value](#func.model.tune.param_range.args.fn_tune.test.param_value) argument to the [model.tune.param_range](#func.model.tune.param_range) function described above in the section *Common Helper Functions /* [Model Tuning Utils] of *this Appendix*.
